Hydras, Ordinals \& Co. \\ A library in Coq of entertaining formal mathematics

### 2021-10-13 doc/JFLA2022/paper.tex
misc additions and fixes

Libraries of formalized mathematics based on proof assistants, such as \coq, Lean, and Isabelle/HOL, are continually growing in size and scope.
For example, the core projects in the Mathematical Components family for Coq amount to more than 160,000 lines of code and 11,000 lemmas~\cite{Nie2020a}, while Lean's Mathlib consists of more than 140,000 lines of code and 34,000 declarations~\cite{Mathlib}. However, after a key mathematical definition or result is added to a library, it must be \emph{documented and maintained}~\cite{vanDoorn2020,alectryonpaper}.

@@

The data types for ordinals in Cantor and Veblen form are essentially defined the same way in Gaia and \Hydras, but they are not identical inside Coq, e.g., due to placement in different modules. Moreover, there are small differences in the implementation of ordinal arithmetic due to different evolutionary paths taken from the ancestral Cantor library.

@@

@@

\begin{table}
\centering
\footnotesize
\begin{tabular}{|l|l|l|l|}
\hline
\textbf{Package} & \textbf{Repository} & \textbf{Spec LOC} & \textbf{Proof LOC}\\
\hline
Pocklington & \texttt{pocklington} & & \\
\hline
Goedel & \texttt{goedel} & & \\
\hline
Gaia & \texttt{gaia} & & \\
\hline
Hydra-battles & \texttt{hydra-battles} & & \\
\hline
Addition-chains & \texttt{hydra-battles} & & \\
\hline
Gaia-hydras & \texttt{hydra-battles} & & \\
\hline
\end{tabular}
\caption{Current numbers of lines of code for \Hydras packages.}
\label{tbl:loc}
\end{table}

@@

\Hydras wants to be a bridge between scientific literature (\emph{e.g.}~\cite{KP82, KS81, schutte} and proof assistant technology. For the mathematician, it may give a concrete view of the mathematical content, not only through full proofs, but also through significative computations: examples, functions associated with constructive proofs, etc. For the \coq learner, it provides a consistent set of examples,
allowing to present and compare various formalization and proving techniques. %It may be also used as a library, through opam packages.
It is also a medium sized library (more than 50K LOC), dependent on various tools and libraries of the \coq ecosystem, which  may be also used to experiment new techniques of continuous maintenance of the code and its documentation.

>>>>

Libraries of formalized mathematics based on proof assistants, such as \coq, Lean, and Isabelle/HOL, are continually growing in size and scope.
For example, the core projects in the Mathematical Components family for Coq amount to more than 160,000 lines of code (LOC) and 11,000 lemmas~\cite{Nie2020a}, while Lean's Mathlib consists of more than 140,000 LOC and 34,000 declarations~\cite{Mathlib}. However, after a key mathematical definition or result is added to a library, it must be \emph{documented and maintained}~\cite{vanDoorn2020,alectryonpaper}.

@@

The data types for ordinals in Cantor and Veblen form are essentially defined the same way in Gaia and \Hydras, but they are not identical inside Coq, e.g., due to placement in different modules. Moreover, there are small differences in the implementation of ordinal arithmetic due to different evolutionary paths taken since divergence from the ancestral Cantor library.

@@

Figure~\ref{fig:genealogy} illustrates the relationships between \Hydras packages.

@@

Table~\ref{tbl:loc} lists the size of each \Hydras package.

\begin{table}
\centering
\footnotesize
\begin{tabular}{|l|l|l|l|l|}
\hline
\textbf{Package} & \textbf{Repository} & \textbf{Commit} & \textbf{Spec LOC} & \textbf{Proof LOC}\\
\hline
Pocklington & \texttt{pocklington} & & & \\
\hline
Goedel & \texttt{goedel} & & & \\
\hline
Gaia & \texttt{gaia} & & & \\
\hline
Hydra-battles & \texttt{hydra-battles} & & & \\
\hline
Addition-chains & \texttt{hydra-battles} & & & \\
\hline
Gaia-hydras & \texttt{hydra-battles} & & & \\
\hline
\end{tabular}
\caption{Current numbers of lines of code for \Hydras packages.}
\label{tbl:loc}
\end{table}

@@

\Hydras wants to be a bridge between scientific literature (\emph{e.g.},~\cite{KP82, KS81, schutte}) and proof assistant technology. For the mathematician, it may give a concrete view of the mathematical content, not only through full proofs, but also through significative computations: examples, functions associated with constructive proofs, etc. For the \coq learner, it provides a consistent set of examples,
allowing to present and compare various formalization and proving techniques. %It may be also used as a library, through opam packages.
It is also a medium sized library (more than 50K LOC), dependent on various tools and libraries of the \coq ecosystem, which  may be also used to experiment new techniques of continuous maintenance of the code and its documentation.
